---
title: "e) Making Runge and Mara Figure 28.3"
author: "Nathan Brouwer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{e) Making Runge and Mara Figure 28.3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Introduction

Figure 28.3 is a 3D surface showing how the overall equilibrium population size (TODO: how defined - all summer individuals?) of the population changes as the amount of high-qualty "source" breeding habitat and high-equality winter habitat varies.  Limitation is indicated by ... TODO


## Preliminaries

Load the redstart package  
TODO add download instructions for development version

```{r}
library(redstart)
```


## Making Figure 28.3 the easy way


As inputs the **runFAC_multi()** function takes a dataframe with two or more parameters varied.  In Figure 28.3 summer (TODO K.x?) and winter (TODOL: K.x?) carrying capacities are each varied systematically from 0 to 1000 (TODO: number of source habitats?).  This range is set up using two functions,  **param_ranges()** and **param_seqs()**.  **param_ranges()** sets the end points of the range, and **param_seqs()** generates the intervening values between the endpoints to actually input into the model.

The defaults for **runFAC_multi()** are set to run param_ranges() and param_seqs() behind the scenes to reproduce Figure 28.3 in Runge and Marra (2004).  The data underlying the figure can therefore be generated just by calling runFAC_multi().  Below, I'll first show how to quickly create the figure using the defaults.  I'll then gradually show how the functions work.

TODO: indicate that breeding carrying capacity is in pairs and winter is in idividuals 

### Making Figure 28.3 automatically

We can generate figure 28.4 by running runFAC_multi() without any arguments.  This may take a minute or two.

```{r}
F28_3 <- runFAC_multi()
```

This produces a list; one of the elements of this list is as dataframe called "multiFAC.out.df.RM" ("multi-FAC output dataframe - Runge and Marra version"), which contains all the equilibirum population sizes which make up the figure.

This dataframe can be plotted with the function **plot_Fig28_3()**.  

First, let's extract out the dataframe with our information
```{r}
F28_3_data <- F28_3$multiFAC.out.df.RM
```

Now build the plot.
TODO: oritation of plot - I have K.wg on the left instead of the right
```{r}
par(mfrow = c(1,1))
plot_Fig28_3(F28_3_data)
```


### Making Figure 28.3: some details on the "easy way"

Below are some insights into what's going on behind the scenes above.  This will set us up for a detailed explanation in the next sections.

The **param_ranges()** functions has a "figure = ..." arguement that allows you to specify a particular figure to generate.  This allows you to set the parameters used for a particular figure in the paper.  The default is for figure 28.3  So whne you call just runFAC_multi() interally what's happneing is this:

```{r}
F28.3.range <- param_ranges(figure = 28.3)
```

Note that K.bc and K.wg both vary from 1 to 1000.
```{r}
head(F28.3.range)
```

Once the endpoints of the ranges are set param_seqs() generates the intervening values.  So, what's happening internally is this:

```{r}
F28.3.seq <- param_seqs(F28.3.range)
```

We can see some of the fixed and also the varying parameters here
```{r}
head(F28.3.seq)
```

These vectors are then transformed internally into a "grid" of all possible combinations using param_grid()
```{r}
F28.3.grid <- param_grid(param.seqs = F28.3.seq)
```

This produces a big matrix!
```{r}
dim(F28.3.grid)
```

We can get a peak at the variation here.
```{r}
head(F28.3.grid[,c(1:5)],12)
```

Once the grid is generated we can plug it into runFAC_multi() the param.grid = ... arguement of runFAC_multi().  This happened automatically above, but we could do this explicitly like this:
```{r, eval = F}
F28_3 <-runFAC_multi(param.grid = F28.3.grid, verbose = F)
```

The output is then plotted as above using plot_Fig28_3().

## Make Figure 28.3: Step-by-step instructions

To make Figure 28.3 *de novo* we need specify more information in **param_ranges()**.  This function takes as its arguements all the parameters in the model, though defaults are set so we only have to change what we are spefically interested in.

You can see the defaults by calling up the help file.  Its unforuntaely a bit dense since the input is a range of the minimum and maxium.
```{r, eval = F}
?param_ranges
```


For generating Figure 28.3 want to work with **K.bc** (carrying capacity of breeding "source" habitat) and **K.wg** (carrying capacity of good winter habitat).  We want to vary each from 1 to 1000 (note that K.bc is in terms of pairs while K.wg is individuals).

```{r}
F28.3.range <- param_ranges(K.bc. = c(1, 1000),
                            K.wg. = c(1, 1000))
```

We can then proceed as  with param_seqs() and param_grid() to generate the necessary dataframe.
```{r}
F28.3.seq  <- param_seqs(F28.3.range)
F28.3.grid <- param_grid(param.seqs = F28.3.seq)
```

We then run the model and plot it like this (not run)
```{r, eval = F}
#run the model
F28_3 <-runFAC_multi(param.grid = F28.3.grid)

#plot
plot_Fig28_3(F28_3$multiFAC.out.df.RM)
```


## Comparing output of redstart to the original 

TODO
