---
title: "Running a full-annual cycle (FAC) model with FACavian"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

The full-annual cycle (FAC) model of Runge and Marra (2004) is a density-dependent (dd) matrix model.  The principal output of a density-dependent model is the **equilibrium population size**.  An important issue with dd models is that you can only determine the equilibrium population size using numerical simulation.  This means that you actually have to plug information -- particularly initial population sizes -- into the model and run it through many iterations and monitor the output until you see that it stabilizes or converges to a stable state. This is in contrast to traditional matrix models (and their cousin, integral projection model) which can be analyzed mathematically without specifying a starting population state or iterating the model (Footnote: assuming they are ergodic; Stott et al 2010 Method E&E)

In the FACavian package, a single run of a FAC is done using the runFAC() function.  The user gives this function three bundles of information:

1. The demographic parameters for the model, including surival rates and  fecundities
1. The habitat carrying capacities
1. The intial population state.  Specifically the abundance and age structure of male and females on the wintering ground in "good" and "poor" habitat.

runFAC() then builds all of the matrices that describe the FAC, plugs in the starting population state, and turns the crank on the model for either a given number of iterations (the default is 300) or, if specified, until a stability criteria is met.

**runFAC()** can be set to run and either keep track of everything that happens during the entire model run (save.ts = T, the default), or just the final time point which (save.ts = F); if the model reach equilibrium, represents that stable population size, stage, and sex structure of the model during each time point.

Examining the entire time series ("ts") of the model is mostly useful for understanding what the model does and for diagnostics purposes, eg to makes sure it is reaching equilbrium.  For most subsequent analyses just the equilibrium state is needed; a typical analysis of an FAC model involves running the model many times while systematically varying one or two key parameters to understand how they impact the equilibrium population state.

In this vignette I'll demonstrate what **runFAC()** does and what the output of a single run the FAC looks like.  In subsequent vignettes we'll using a wrapper function, **runFAC_multi()** to run the model repeatedly to equilibirum while varying parameters such as carrying capacity.  When doing simple explorations of how the function works I'll set return.output = F so that the full output dataframe isn't printed out into the R terminal.


## Preliminaries

### Getting the FACavian package

The most recent version of FACavian is available from github and can be downloaded directly using hte devtools package
```{r eval=FALSE}
#load devtools
library(devtools)

#download FACavian from github
devtools::install_github("brouwern/FACavian")


```


Load FACavian into your current R sessions
```{r}
library(FACavian)
```



### Getting parameters for a model run

Default values are set in all of FACavian's functions so you can easily explore them without worrying about setting the underlying demographic parameters.  For a preliminary run parameters will be called automatically behind the scenes using a function called param_set(), which has defaults as givenn in Runge and Marra (2004).  

You can look at the parameters by calling the param_set() function.  Later we'll use param_set() to tweak a few parameters.

Counting the 3 carrying capacities there are 30 parameters in a single run of the rnFAC.
```{r}
length(param_set())
```


The first 10 parameters are:
```{r}
param_set()[1:10]
```


Additionally, there are four intial population size values for males and females in the two types of winter habitats.  Defaults are set for these using the arguement "Ninint = c(....) within runFAC.

## Running a FAC


When you call runFAC() a list is created.  This list contains everything you put into runFAC() (the intial parameters) and everyting the model produced. 

We can run the model and save its output to an object
```{r}
my.fac <- runFAC()
```

The output is a list
```{r}
is(my.fac)
```

This list has 8 elements.
```{r}
length(my.fac)
```

This is dense object - some of the elements in the list are list themselves!  For example, the element "param.matrices" is a list which contains all of the transition matrices for the modle

```{r}
is(my.fac$param.matrices)
```

To look at the names of all the elements of the list we can use the str() command to look at the output.   The "1" tells str() to just give us the names of the elements inthe list.

```{r}
str(my.fac,1)
```


The elements of the list are

1. **FAC.out.RM**: a dataframe containing the outputs from the original Runge and Mara FAC model (rmFAC)
1. **FAC.out.IB**: a daframe containing the output from a slightly modified version of the model which uses an individual-based modeling (IBM) submodel (ibFAC) for some calculations.
1. **FAC.eq.state.RM**: a vector with the state of the original rmFAC when the model reaches equilibrium
1.  **FAC.eq.state.IB**: a vector with the state of the  ibFAC when the model reaches equilibrium.
1. **W.mg.diagnostic.df**:  TODO()
1. **param.matrices**: All of the parameterized transition matrices describing the model
1. **params.initial**: initial parameters used to set up the transition matrices
1. **use.IBM**: indicater (TRUE/FALSE) if the IBM sub-model was run.


The main output of runFAC() is "FAC.out.RM" and "FAC.eq.state.RM." When many runs of the model are required we can just save the equilibrium state to improve performace of the model. 



### A basic run of the FAC

We'll call runFAC() with the argument return.output set to FALSE to supress the output of full time series.  Also set diagnostic.plot = T to see plot of the time series.

To be clear about what the options do, if I ran just this
```{r, eval=F}
runFAC()
```

I'd run then model and then get a big list with all the data output to my R terminal.  

In contrast, the code below will run the model to equilibrium but won't output anything.
```{r}
runFAC(return.output = F)
```



We can see how the model progresses to equilbibrium by setting diagnostic.plot = T.
```{r, eval = T}
runFAC(return.output = F, 
       diagnostic.plot = T)
```



Our diagnostic plot has three panels.  On the left hand size is a graph of total winter population size (tot.W; black) and total breeding poulation size (tot.B, red).  Winter population size is larger because birds die during the winter and spring migraiton.  The middle panel fo the graph shows a simple estimate of population growth rate,  calcualted simply as winter population size in year t+1 divided by size in year t.  When growth stabilizes at 1.0, the model is at equilibrium.


The default behavior of runFAC() is to monitor the output for equilibrium.  If you want to make sure your model is at equilibrium you can turn off the default equilibirum monitoring using the check.eq arguement; the model will run for the default of 350 iterations.

```{r, eval=T}
runFAC(return.output = F,
       check.eq = FALSE,
       diagnostic.plot = T)
```

You can also specify a specific number of interations
```{r, eval=FALSE}
###: NOTE: plot returned here is off; doesn't include zero
runFAC(return.output = F,
       iterations = 20,
       diagnostic.plot = T
       )
```


We can also have it run long by making the equilibrium checker -- which checks the time series for an asymptote -- stricter.  This can be done by increase the "equilibrium tolerance", eq.tol.  The model will still check if the time series has reached an asymptote (check.eq = TRUE) but will be stricter about judging equilibrium.

```{r, eval=F}
runFAC(return.output = F,
       eq.to = 10,
       diagnostic.plot = T)
```


After we run a model we can also create the plot after the face

```{r, eval=F}
my.fac <- runFAC(iterations = 150)
plot_runFAC(my.fac$FAC.out.RM)

```



###

The main output of runFAC() is a dataframe containing the status of the population at each timestep.  This time series contains information about how everything in the model jchanges over tiem.  We can access this dataframe by removing "return.output = F".  The default of runFAC() is returnet.output = T, but we've set it to false so we don't have to see all the crazy  output.   

The dataframe is 45 columns wide and has the state variables for each step in the model, as well as summary stats (eg sums of several column) such as the total breeding (tot.B) and  winter (tot.W) populations.

runFAC() outputs a list with many things, and the timeseres dataframe is contained in an item called "FAC.out.RM"


If you run this code you'll get a big dataframe (not iteractions is set to a small number to make it run fast)
```{r, eval = F}
runFAC(iterations = 5)["FAC.out.RM"]
```

We can look at the data used in the plotting diagnostic like this

First, run the model and assign the FAC.out.RM" part to an object (note the double square brackets)
```{r}
my.FAC<- runFAC()[["FAC.out.RM"]]
```

Now look at some focal columns

```{r}
head(my.FAC[,c("tot.B","tot.W","lambda")])
```


Each row in this dataframe is from one iteraction of the model.  NAs are very common in this dataframe - they indicate that the model didn't run long enough to assign values to that iteration.
```{r}
tail(my.FAC[,c("tot.B","tot.W","lambda")])
```



### Setting parameters

### Returning just the equilibrium state



```{r}
my.fac <- runFAC(save.ts = F)
```



```{r}
my.fac$FAC.eq.state.RM
```


TODO(): simplfy output of equilibrium output



