---
title: "z) Building and rebuilding vignettes"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval  = F,
  comment = "#>"
)
```

## Introduction

This code is used to build, re-build, and trouble shoot the building of the package, including 

1. Building/compiling diffrent components
1. Building the pkgdown website

These are mostly commands that can be launched from within RStudio via the GUI, except pkgdown::build_site()

## devtools functions

```{r}
library(devtools)
```


### Clean vignettes

"This uses a fairly rudimentary algorithm where any files in ‘inst/doc’ with a name that exists in ‘vignettes’ are removed."

```{r, eval = F}
clean_vignettes() 
```

### Build vignettes

"Builds package vignettes using the same algorithm that R CMD build does"

```{r, eval = F}
devtools::build_vignettes()
```

### Build documentation

document()

```{r, eval = F}
devtools::document(roclets=c('rd', 'collate', 'namespace'))
```

### Load all

```{r, eval = F}
devtools::load_all(".")

```


### Check

```{r, eval = F}
devtools::check()
```

check_failures

## pckgdown functions

### Build package down website

Single function 
```{r, eval = F}
pkgdown::build_site()
```


#### Build piece by piece

pkgdown::clean_site (pkg = ".")

```{r,eval = F}
pkgdown::init_site(pkg = ".")
```

```{r, eval = F}
pkgdown::build_reference(pkg = ".")
```

```{r, eval = F}
build_news()
```


```{r, eval = F}
pkgdown::build_articles(pkg = ".")
```


```{r, eval = F}
pkgdown::build_home(pkg = ".")
```



## Explore dependencies

```{r}
report1 <- CreatePackageReport(pkg_name = "FACavian")
```



## Rcmd.exe
"Clean and rebuild"
==> Rcmd.exe INSTALL --preclean --no-multiarch --with-keep.source mammalsmilk

"Install restart"
==> Rcmd.exe INSTALL --no-multiarch --with-keep.source mammalsmilk





